### 2021-11-15 input/proposal.tex
mostly about algo text

\begin{definition}[\label{def:purge}Message purge]
  A system purges a message $P(m)$ when every \node $q$ that delivered
  $m$ eventually delivers another message $m'$ never followed by the
  delivery of the former message $m$: $d_q(m) \implies \eventually
  (d_q(m) \rightarrow d_q(m'))$.
\end{definition}

\begin{theorem}[\label{theo:dcp}EFB$^+$: EFB $\wedge$ (Purge $\rightarrow$ EFB) $\implies$
    \underline{D}ynamic \underline{c}onsistent
    \underline{p}artitioning (DCP)]
%
  When a \process can \texttt{Add} and \texttt{Del}ete its partition,
  to ensure consistent partitioning, eventual forwarding of best
  (Theorem~\ref{theo:efb}) additionally requires
\begin{inparaenum}[(i)]
\item eventual purging of stale messages ($S(m) \implies P(m)$), then
\item the eventual forwarding of best (Definition~\ref{theo:efb}).
\end{inparaenum}
\end{theorem}

\begin{proof}
  %% To reach consistent partitioning, the last message delivered by
  %% every process is not stale and comes from its respective best
  %% source: $d(m)$
  Assuming EFB as default propagation behavior where \processes
  deliver and forward the best sum of weights they received. We must
  prove that despite stale messages, the last delivery of every
  \process is the message that followed the shortest path to its
  closest source.
  \begin{asparadesc}
  \item [EFB without Purge:] From Figure~\ref{fig:problemD}, a
    \process may receive messages in such an order ($d_b(\alpha_d^1)
    \rightarrow r_b(\delta_a) \rightarrow d_b(\delta_d)$) that it does
    not forward messages required by other \processes
    ($\delta_a$). The last delivery of some nodes ($c$ and $d$) is
    wrong ($\alpha_a^3$ instead of $\delta_a$).
    % \Processes never
    % reach consistent partitioning.
    %
    %% Figure~\ref{fig:proofA} depicts the generic case where a \process
    %% $b$ precludes itself and subsequent \processes such as $c$ of
    %% delivering the message from their respective best source
    %% $a$. \Process $b$ received and delivered a message from a
    %% partition that has since then been deleted but with a better sum
    %% of weights than newly received ones ($\alpha_d^y <
    %% \alpha_a^{x'}$). It forever discards such a message
    %% $\alpha_a^{x'}$.
  \item [Purge not followed by EFB:]
    %% also transiting messages
    Assuming that every \process removes control information as soon
    as it becomes stale ($S(\alpha_a) \implies d_b(\alpha_a^x)
    \rightarrow d_b(\delta_a)$) and cannot deliver stale messages
    anymore ($S(\alpha_a^x) \implies d_c(m) \rightarrow \ldots
    \rightarrow d_c(m' \not\in \alpha_a)$), staleness does not impair
    up-to-date messages propagation. However, even in such settings,
    depending on receipt order ($d_b(\alpha_a^{x}) \rightarrow
    r_b(\alpha_c^{y : y > x}) \rightarrow d_b(\delta_b)$), a \process
    may never receive again, hence deliver and forward a message that
    was received and discarded before ($\alpha_c^{y}$). When such a
    message contains its best up-to-date partition, a \process may
    never end up in its best partition.
    %    Therefore, \processes do not reach
    % consistent partitioning.
  \item [Purge then EFB:] Assuming the eventual and definitive removal
    of stale messages at each \process ($S(\alpha_a) \implies
    (d_b(\alpha_a) \implies \eventually d_b(\alpha_a) \rightarrow
    d_b(m') \not\rightarrow d_b(\alpha_a))$); and assuming that every
    removal eventually triggers the forwarding of best delivered
    messages ($d_c(\delta_x) \implies \eventually f_d(\alpha_y)$),
    \processes may deliver other stale messages that will be 
    eventually removed to never be delivered again. In the worst case,
    all \processes deliver and remove all stale messages to ultimately
    deliver and forward only up-to-date messages ($\eventually d_c(m')
    \implies \neg S(m')$). As in Theorem~\ref{theo:efb}, they
    eventually deliver and forward their best up-to-date message.

@@

\begin{theorem}[\label{theo:threephase}Three-phase purge]
  Three phases are sufficient to eventually purge the system from
  stale control information:
  \begin{inparaenum}[(i)]
  \item propagation of delete notifications,
  \item detection of possibly blocking conditions, and
  \item propagation of delete notifications that were possibly blocked.
  \end{inparaenum}
\end{theorem}

@@

Algorithm~\ref{algo:ascast} provides the few instructions of \NAME
that implement three-phase purge and eventual forwarding of best to
enable dynamic consistent partitioning. For the sake of simplicity, it
does not handle multiple sessions and associated optimisation such as
data structure sharing and message packing.

@@

As stated in Theorem~\ref{theo:dcp}, every \process must purge stale
messages.  To identify staleness, each \process maintains a vector of
versions that associates the respective known local counter of each
known source, or has-been source. Its size increases linearly with the
number of sources that the \process has ever known, which is the
number of \processes in the system $\mathcal{O}(V)$ in the worst case.
Nevertheless, following the principles of scoped broadcast, we expect
that every \process only acknowledges a small subset of sources.
Using such a vector, every message $m$ carries its source $s$ along
with its version $c$ that we subscript $m_{s, c}$.  Each \process
delivers and forwards
\begin{inparaenum}[(i)]
\item only newest messages (see Line~\ref{line:ascast_version}) that
\item improve their best known partition (see
  Line~\ref{line:ascast_better}).
\end{inparaenum}
This vector of versions constitutes a summary of local deliveries. It
allows a \process to detect inconsistencies in message delivery: the
parent delivered a stale message (see Line~\ref{line:ascast_version}),
or the parent delivered a worse message (see
Line~\ref{line:detectA}). % This may happen in stage II of
% Figure~\ref{fig:proofB} between $b$ and $c$.

A \process ($c$) detecting an inconsistency disseminates a $\Delta$
message. Such a $\Delta$ must not modify any vector of versions since
the partition of the corresponding $\alpha$ may still exist.
Therefore, it additionally includes and maintains a path $\pi$
initialized to $\alpha$'s one. Such a $\Delta_{s, c}^{\pi}$
transitively tracks and purges $\alpha$ messages that were forwarded
by the detecting \process. The path $\pi$ also enables a quick stop of
looping messages and in turns ensures termination. However, carrying
paths in messages is costly. In the worst case, a message piggybacks
the identity of all \processes in the system. Fortunately, message
paths and \NAME synergyze well with each other, for paths tend to be
smaller as the number of sources in the system increases. Contrarily
to $\alpha$ and $\Delta$ messages, $\delta_{s, c}$ messages do not
carry a path, and are therefore more efficient. They
\begin{inparaenum}[(i)]
\item solely rely on versions to operate, and
\item use the whole dissemination graph to propagate in the system, as
  opposed to $\Delta$ that must follow the dissemination tree of the
  corresponding $\alpha$.
\end{inparaenum}

As stated in Theorem~\ref{theo:dcp}, dynamic consistent partitioning
not only requires eventual purging of stale messages, but also the
retrieval of the best up-to-date partitions.  In that regard, both
$\delta$ and $\Delta$ messages have dual use: since they already reach
the borders of their partition by design, they trigger a competition
when reaching bordering \processes (see
Line~\ref{line:ascast_compete}).
%% This simply consists in sending the
%% current partition through the communication link from which the
%% \process received the $\delta$. Upon receipt of this answer,
%% \processes act normally by propagating their changes when they
%% improve.

@@

Algorithm~\ref{algo:edges} enables dynamic consistent partitioning in
dynamic networks where \processes can join, leave, or crash at any
time. Removing a \process is equivalent to removing all its edges, and
adding a \process is equivalent to add as many edges as necessary.
Adding an edge between two \processes may only improve the shortest
path of one of these \processes. Therefore, it triggers a competition
between the two \processes only. If one or the other \process
improves, the propagation of $\alpha$ messages operates normally.
Removing an edge between two \processes may invalidate the shortest
path of one of involved \processes plus downstream \processes. As a
side effect, removing an edge may also impair the propagation of a
partition delete. To implement this behavior, \NAME uses its
implementation of $\Delta$ messages. This prove costly but enables
\NAME even in dynamic networks subject to physical partitioning.

>>>>

\begin{definition}[\label{def:purge}Message purge]
  A system purges a message \TODO{$P(m)$} when every \node $q$ that
  delivered $m$ eventually delivers another message $m'$ never
  followed by the delivery of the former message $m$: $d_q(m) \implies
  \eventually (d_q(m) \rightarrow d_q(m')) \TODO{\wedge}$.
\end{definition}

\begin{theorem}[\label{theo:dcp}EFB$^+$: EFB $\wedge$ (Purge $\rightarrow$ EFB) $\implies$
    \underline{D}ynamic \underline{c}onsistent
    \underline{p}artitioning (DCP)]
%
  When a \process can \texttt{Add} and \texttt{Del}ete its partition,
  to ensure consistent partitioning, eventual forwarding of best
  (Theorem~\ref{theo:efb}) additionally requires
\begin{inparaenum}[(i)]
\item eventual purging of stale messages ($S(m) \implies P(m)$), then
\item the eventual forwarding of best (Theorem~\ref{theo:efb}).
\end{inparaenum}
\end{theorem}

\begin{proof}
  %% To reach consistent partitioning, the last message delivered by
  %% every process is not stale and comes from its respective best
  %% source: $d(m)$
  Assuming EFB as default propagation behavior where \processes
  deliver and forward the best sum of weights they received. We must
  prove that despite stale messages, the last delivery of every
  \process is the message that followed the shortest path to its
  closest source.
  \begin{asparadesc}
  \item [EFB without Purge:] A \process may receive messages in such
    an order (\eg in Figure~\ref{fig:problemC}: $d_b(\alpha_c^1)
    \rightarrow r_b(\delta_a) \rightarrow r_b(\delta_a)$) that it does
    not forward messages required by other \processes ($\delta_a$ for
    $c$). The last delivery of some nodes may be wrong (\eg in
    Figure~\ref{fig:problemD}: $\alpha_a^3$ instead of $\delta_a$).
    % \Processes never
    % reach consistent partitioning.
    %
    %% Figure~\ref{fig:proofA} depicts the generic case where a \process
    %% $b$ precludes itself and subsequent \processes such as $c$ of
    %% delivering the message from their respective best source
    %% $a$. \Process $b$ received and delivered a message from a
    %% partition that has since then been deleted but with a better sum
    %% of weights than newly received ones ($\alpha_d^y <
    %% \alpha_a^{x'}$). It forever discards such a message
    %% $\alpha_a^{x'}$.
  \item [Purge not followed by EFB:]
    %% also transiting messages
    Assuming that every \process removes control information as soon
    as it becomes stale ($S(\alpha_a) \implies d_b(\alpha_a^x)
    \rightarrow d_b(\delta_a)$) and cannot deliver stale messages
    anymore ($S(\alpha_a^x) \implies d_c(m) \rightarrow \ldots
    \rightarrow d_c(m' \not\in \alpha_a)$), staleness does not impair
    up-to-date messages propagation. However, even in such settings,
    depending on receipt order ($d_b(\alpha_a^{x}) \rightarrow
    r_b(\alpha_c^{y : y > x}) \rightarrow d_b(\delta_b)$), a \process
    may never receive again, hence deliver and forward a message that
    was received and discarded before ($\alpha_c^{y}$). When such a
    message contains its best up-to-date partition, a \process may
    never end up in its best partition.
    %    Therefore, \processes do not reach
    % consistent partitioning.
  \item [Purge then EFB:] Assuming the eventual and definitive removal
    of stale messages at each \process ($S(\alpha_a) \implies
    (d_b(\alpha_a) \implies \eventually d_b(\alpha_a) \rightarrow
    d_b(m') \not\rightarrow d_b(\alpha_a))$); and assuming that every
    removal eventually triggers the forwarding of best delivered
    messages ($d_c(\delta_x) \implies \eventually f_d(\alpha_y)$),
    \processes may deliver other stale messages that will be 
    eventually removed to never be delivered again. In the worst case,
    all \processes deliver and remove all stale messages to ultimately
    deliver and forward only up-to-date messages ($\eventually d_c(m')
    \implies \neg S(m')$). As in Theorem~\ref{theo:efb}, they
    eventually deliver and forward their best up-to-date message.

@@

\begin{theorem}[\label{theo:threephase}Three-phase purge]
  Three phases are sufficient to eventually purge the system from
  stale control information:
  \begin{inparaenum}[(i)]
  \item propagation of delete notifications,
  \item detection of possibly blocked deletion, and
  \item propagation of delete notifications that were possibly blocked.
  \end{inparaenum}
\end{theorem}

@@

Algorithm~\ref{algo:ascast} provides the instructions of \NAME that
implement three-phase purge and eventual forwarding of best to enable
dynamic consistent partitioning. For the sake of simplicity, it does
not handle multiple sessions nor optimisation such as message packing.

@@

As stated in Theorem~\ref{theo:dcp}, every \process must purge stale
messages.  To identify staleness, each \process maintains a vector of
versions that associates the respective known local counter of each
known source, or has-been source. Its size increases linearly with the
number of sources that the \process has ever known, which is the
number of \processes in the system $\mathcal{O}(V)$ in the worst case.
Nevertheless, following the principles of scoped broadcast, we expect
that every \process only acknowledges a small subset of sources.  In
addition, each message carries the identifier and counter of each
\node that forwarded it. Again, following the principles of scoped
broadcast, we expect that versioned paths remain small. Using these
data structures, each \process delivers and forwards only new messages
that improve their best known partition (see
Line~\ref{line:ascast_improve}).

This vector of versions constitutes a summary of known progress of
other \nodes. It discards stale messages and it enables detecting
possible inconsistencies in message delivery: the current best
partition comes from a neighbor that just delivered a stale message
(see Line~\ref{line:ascast_detect}). Figure~\ref{fig:problem} shows
that such a behavior may hide a delete notification. The detecting
\process must act to reflect this possible change: all \processes
whose last message comes from the detecting \process must remove it
and accept the competition of other \processes. Since this behavior is
similar to that of source deletion, we implement it in a similar way.

A \process that removes its self-appointed status of source, or
detects a possible inconsistency, increments its local counter and
disseminates a delete notification message $\delta_{s, c}$
piggybacking its identifier $s$ and counter $c$. Any \process that
receives such delete notification deliver and forward it if their best
message contains the has-been source or detector associated to a lower
counter (see Line~\ref{line:ascast_delete}). This allows $\delta$
messages to quickly reach every downstream \process.

As stated in Theorem~\ref{theo:dcp}, dynamic consistent partitioning
not only requires eventual purging of stale messages, but also the
retrieval of the best up-to-date partitions.  In that regard, $\delta$
messages have a dual use: since they already reach the borders of
their partition by design, they trigger a competition when reaching
bordering.  When a \process receives a delete notification from a
neighbor $q$ but does not deliver it, it assumes that the just
received $\delta$ message purged possibly stale messages, creating a
gap that its own best partition could help fill. Consequently, it
sends to $q$ its own best $\alpha$ message (see
Line~\ref{line:ascast_echo}).

@@

Algorithm~\ref{algo:edges} enables dynamic consistent partitioning in
dynamic networks where \processes can join, leave, or crash at any
time. Removing a \process is equivalent to removing all its edges, and
adding a \process is equivalent to add as many edges as necessary.
Adding an edge between two \processes may only improve the shortest
path of one of these \processes. Therefore, it triggers a competition
between the two \processes. If one or the other \process improves, the
propagation of $\alpha$ messages operates normally.  Removing an edge
between two \processes may invalidate the shortest path of one of
involved \processes plus downstream \processes. As a side effect,
removing an edge may also impair the propagation of a partition
delete. To implement this behavior, \NAME uses its implementation of
$\delta$ messages. This enables \NAME even in dynamic networks subject
to physical partitioning.
